### 2016-09-07 syntax.tex
updates to syntax.tex

\noindent The definitions above can be lifted to the random-oracle
model by giving $\Rep, \Qry$ access to a random oracle. In that
case, the probabilities defining the error probability and error
rate are also taken over choice of the random oracle.

We will sometimes say that a data structure is
$(1-\epsilon)$-correct, meaning it has error probability~$\epsilon$.
Note that $1-\epsilon$ is a worst-case lower bound on the
correctness of the data structure. \tsnote{Don't know if this will be necessary.}

@@

\def\bin{{\sf Bin}}
\heading{``Trivial'' attacks, and error probability vs.\ error
rate.} \jnote{Not sure this belongs here. I'd rather move any
discussion of ``attacks'' (trivial or not) to later. But if we just
want to highlight the difference between error and error rate we
could do that here.} Let~$\Pi$ be a scheme that is
$(1-\epsilon)$-correct, and consider an attack that asks random
queries $q_1,q_2,\ldots,q_T$, i.e., uniform samples from~$\calQ$.
Let $\mathrm{Err}_i$ be a random variable indicating the event
$\Qry(\pubaux,\privaux,q_i)\neq q_i(S)$. The
$(1-\epsilon)$-correctness of~$\Pi$ guarantees that the probability
that $\mathrm{Err}_i=1$ is at most~$\epsilon$, and it is tempting to
model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials with success
probability~$\epsilon$. However, a moment's reflection shows that
this is not the case.  Concretely, consider the following scheme.
Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$. For the
queries, let $\calQ = \{q_x\colon \calS \to \bits \,|\, \forall x
\in \univ\}$ where $\forall S \in \calS$ the predicate $q_x(S)=1
\Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random element~$s
\in S$, and return $M = S \setminus \{s\}$ and $\privaux =
\emptystring$.  To respond to queries,
let~$\Qry(\pubaux,\privaux,q_x)=1$ iff $x \in M$.  It is easy to see
that this scheme is $1/n$-correct; for any fixed~$S$, there are
exactly~$n$ queries~$q_x$ that have probability~$1/n$  (over the
coins of~$\Rep$) of causing a correctness error, and the remaining
$|\univ|-n$ queries have probability zero of causing an error.
Moreover, once the coins of~$\Rep$ are fixed, there is exactly
\emph{one} query that will cause an error.  Thus for random queries
from~$Q$, for all~$i\in[T]$ we have
$\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

This example is contrived, but it highlights the important
difference between error probability and error rate.  In particular,
the error probability~$\epsilon$ is measured only over the coins
of~$\Rep$, for a fixed~$S$ and query~$q$. Once the coins of~$\Rep$
are fixed, it is not straightforward how to connect~$\epsilon$ to
events that depend on other sources of randomness. \tsnote{Not sure
if there's more to say in this section.}

\heading{Data structures for set-multiplicity queries.} Our
particular focus will be on data structures that support
multiplicity queries, of which set-membership queries are a special
case.
\begin{definition} \rm
A \emph{set-multiplicity data structure} is a data structure with
$\Gamma=\mathbb{N}$ and $\calQ=\{q_x\}_{x \in \univ}$, where
$q_x(S)$ is defined to be the multiplicity of~$x$ in the
multiset~$S$. \hfill\dqed
\end{definition}

We illustrate our notation by casting Bloom filters as a particular
example of a set-multiplicity data structure. \jnote{Only handles
sets, not multisets.} Let $\Gamma=\bits$, and fix functions
$k(\cdot)$ and $m(\cdot)$ parameterizing the scheme. Algorithm
$\Rep$, on input a set~$S$, computes $k=k(|S|)$ and $m=m(|S|)$ and
then chooses $k$ functions $h_1, \ldots, h_k \in \mathcal{H}$ for
some function family $\mathcal{H}=\{h \colon \univ \to
\{1,2,\ldots,m\}\}$. \jnote{Note depends on $m$. It might be easier
(especially since we talk about error) to already let the hash
functions be independent random oracles.} It then computes an
$m$-bit array~$M$ (initialized to 0 everywhere) by setting
$M[h_i(x)]=1$ for all $i\in [k]$ and $x \in S$.  Finally, $\Rep$
returns $\privaux=\emptystring$ and $\pubaux = M$ as the
representation.
%
Algorithm~$\Qry$ is defined so that, on input
$(\pubaux,\privaux,q_x)$, it returns the minimum of the values
$M[h_1(x)],\ldots,M[h_k(x)]$. \jnote{How does $\Qry$ know the hash
functions?}

As a simple extension, consider $\Gamma=\mathbb{N}$ and modify the
basic Bloom filter so that it stores counters, rather than bits, at
each position in an array.  In particular, for each~$x \in S$ the
$\Rep$~algorithm increments the counters stored at positions
$h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of
counters as~$M$.  The $\Qry$-algorithm remains the same. Such a
construction gives a static version of a counting Bloom
filter~\cite{xxx}.  We will capture dynamic versions, as well as
count-min sketches~\cite{xxx}, scalable Bloom filters~\cite{xxx},
etc., in a later section.\tsnote{If we ultimately do!}

@@

By setting $\Sigma=\bits^L$ for a specified $L>0$, we can capture
the so-called ``garbled Bloom filter'' due to Dong, Chen and
Wen~\cite{xxx}. \jnote{Details have to be filled in.} This stores
xor-shares of bitstring~$x \in S$ at positions
$h_1(x),h_2(x),\ldots,h_k(x)$ in an array. Similarly, the
cuckoo-hashing construction from NY, which (loosely speaking) stores
a hash $g(x)$ at positions determined by
$h_1(x),h_2(x),\ldots,h_k(x)$, is captured when $L=|g(x)|$.
\tsnote{Check the correctness of this claim.}

>>>>

%We will sometimes say that a data structure is
%$(1-\epsilon)$-correct, meaning it has error probability~$\epsilon$.
%Note that $1-\epsilon$ is a worst-case lower bound on the
%correctness of the data structure. \tsnote{Don't know if this will be necessary.}

These definitions, and those that follow, can be lifted to the random-oracle
model by giving $\Rep, \Qry$ access to a random oracle. In that
case, the probabilities defining the error probability and error
rate are also taken over choice of the random oracle.

@@

\def\bin{{\sf Bin}}
\heading{Error probability vs.\ error rate. }
We pause to highlight an important
difference between error probability and error rate.  In particular,
the error probability~$\epsilon$ is measured only over the coins
of~$\Rep$, for a fixed~$S$ and query~$q$. Once the coins of~$\Rep$
are fixed, it is not straightforward how to connect~$\epsilon$ to
events that depend on other sources of randomness.   We illustrate how
this surfaces with the following example.

Let~$\Pi$ be a data structure that has error-probability $\epsilon$,
and consider a correctness-violating attack that asks random
queries $q_1,q_2,\ldots,q_T$, i.e., uniform samples from~$\calQ$.
Let $\mathrm{Err}_i$ be a random variable indicating the event
$\Qry(\pubaux,\privaux,q_i)\neq q_i(S)$.  We know that the probability
that $\mathrm{Err}_i=1$ is at most~$\epsilon$, and it is tempting to
model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials with success
probability~$\epsilon$. However, a moment's reflection shows that
this is not the case.  Concretely, consider the following scheme.
Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$. For the
queries, let $\calQ = \{q_x\colon \calS \to \bits \,|\, \forall x
\in \univ\}$ where $\forall S \in \calS$ the predicate $q_x(S)=1
\Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random element~$s
\in S$, and return $M = S \setminus \{s\}$ and $\privaux =
\emptystring$.  To respond to queries,
let~$\Qry(\pubaux,\privaux,q_x)=1$ iff $x \in M$.  It is easy to see
that this scheme is $1/n$-correct; for any fixed~$S$, there are
exactly~$n$ queries~$q_x$ that have probability~$1/n$  (over the
coins of~$\Rep$) of causing a correctness error, and the remaining
$|\univ|-n$ queries have probability zero of causing an error.
Moreover, once the coins of~$\Rep$ are fixed, there is exactly
\emph{one} query that will cause an error.  Thus for random queries
from~$Q$, for all~$i\in[T]$ we have
$\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

\heading{Data structures for set-multiplicity queries.} Our
particular focus will be on data structures that support
multiplicity queries.

\begin{definition} \rm
A \emph{set-multiplicity data structure} is a data structure with
$\Gamma=\mathbb{N}$ and $\calQ=\{q_x\}_{x \in \univ}$, where
$q_x(S)$ is defined to be the multiplicity of~$x$ in the
multiset~$S$. \hfill\dqed
\end{definition}

The classical Bloom filter is a special case of a set-multiplicity
data structure, as set-membership query is a type of
set-multiplicity query.  Specifically, 
%\jnote{Only handles sets, not multisets.} 
let $\Gamma=\bits$, and fix functions
$k(\cdot)$ and $m(\cdot)$ parameterizing the scheme. Algorithm
$\Rep$, on input a set~$S$, computes $k=k(|S|)$ and $m=m(|S|)$ and
then chooses $k$ functions $h_1, \ldots, h_k \in \mathcal{H}$ for
some function family $\mathcal{H}=\{h \colon \univ \to
\{1,2,\ldots,m\}\}$. \jnote{Note depends on $m$. It might be easier
(especially since we talk about error) to already let the hash
functions be independent random oracles.} \tsnote{I don't see the
problem.  $\Rep$ knows~$m$, so any function (whose range admits at
least~$m$ values) can be effectively treated as having range~$[m]$.
But I added a sentence at the end of this paragraph, too.  See if that
helps.} 
It then computes an $m$-bit array~$M$ (initialized to 0 everywhere) by setting
$M[h_i(x)]=1$ for all $i\in [k]$ and $x \in S$.  Finally, $\Rep$
returns $\privaux=\emptystring$ and $\pubaux = \langle M,h_1,h_2,\ldots,h_k\rangle$ as the
representation.
%
Algorithm~$\Qry$ is defined so that, on input
$(\pubaux,\privaux,q_x)$, it returns the minimum of the values
$M[h_1(x)],\ldots,M[h_k(x)]$.  We note that the standard Bloom filter
analysis treats $h_1,h_2,\ldots,h_k$ as independent random functions,
which we may capture in the ROM with some simple adjustments to the above.

As a simple extension, consider $\Gamma=\mathbb{N}$ and modify the
classical Bloom filter so that it stores counters, rather than bits, at
each position in an array.  In particular, for each~$x \in S$ the
$\Rep$~algorithm increments the counters stored at positions
$h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of
counters as~$M$.  The $\Qry$-algorithm remains the same. Such a
construction gives a static version of a counting Bloom
filter~\cite{xxx}.  We will capture dynamic versions, as well as
count-min sketches~\cite{xxx}, scalable Bloom filters~\cite{xxx},
etc., in future work.\tsnote{?}

@@

By setting $\Sigma=\bits^L$ for a specified $L>0$, we can capture
the so-called ``garbled Bloom filter'' due to Dong, Chen and
Wen~\cite{xxx}. \jnote{Details have to be filled in.} This stores
xor-shares of bitstring~$x \in S$ at positions
$h_1(x),h_2(x),\ldots,h_k(x)$ in an array. Similarly, the
cuckoo-hashing construction from NY, which (loosely speaking) stores
a hash $g(x)$ at positions determined by
$h_1(x),h_2(x),\ldots,h_k(x)$, is captured when $L=|g(x)|$.
\todo{Check the correctness of this claim.}
